Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View

### 2020-07-15 Manuscript/sec4-experiments.tex
更新了一些问题

Fig~\ref{fig:exp_absolute_training_time} compares the training time per epoch for each GNN model, the ranking is $\boldsymbol{GaAN} \gg \boldsymbol{GAT} > \boldsymbol{GGNN} > \boldsymbol{GCN}$.
Its time-consuming ranking is consistent with the complexity analysis because the number of edges in the graph generally far exceeds the number of nodes, 
the GAT algorithm with higher edge computing complexity is more time-consuming than the algorithm GGNN with high vertex computational complexity. At the same time, 
Fig~\ref{fig:exp_absolute_training_time} shows that the training time of individual epochs is abnormally high, which is mainly caused by profiling overhead and the GC pause of the Python interpreter.
This phenomenon confirms the necessity of removing abnormal epochs.

@@

\begin{figure}
	\centering
    \subfloat[GCN\label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gcn}]{\includegraphics[height=4cm]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gcn.png}}\\
    \subfloat[GGNN\label{fig:exp_hyperparameter_on_vertex_edge_phase_time_ggnn}]{\includegraphics[height=4cm]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_ggnn.png}}\\
    \subfloat[GAT\label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat}]{\includegraphics[height=4cm]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gat.png}}\\
    \subfloat[GaAN\label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gaan}]{\includegraphics[height=4cm]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gaan.png}}
    \caption{The influence of hyperparameters on GNN vertex/edge calculation time}
	\label{fig:exp_hyperparameter_on_vertex_edge_phase_time}
\end{figure}

@@

Experiments also show that \textbf{the average degree of the dataset affects the time-consuming proportion of vertex/edge calculation}. 
Our fixed graph has 50k vertices, and the R-MAT generator is used to generate a random graph with an average degree between 10 and 100.
We measure the change of the time-consuming proportion of the vertex/edge calculation in each GNN model varies with the average degree of the graph,
as shown in Fig~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}. The time calculation of the edge increases linearly with the increase of the average degree,
\textbf{The time calculation of the edge calculation dominates the entire calculation time in most cases}. 
Only when the vertex calculation complexity is very high and the average degree is very low, the vertex calculation time can overtake the edge calculation time.
Therefore, the focus of \textbf{GNN training optimization should improve the efficiency of edge calculation}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{figs/illustration/steps_in_edge_calculation.png}
    \caption{Step decomposition of edge calculation}
    \label{fig:steps_in_edge_calculation}
\end{figure}

@@

Fig~\ref{fig:exp_sampling_batch_train_time} shows the change of the training time of each batch on the `amc` and `fli` datasets with the batch size after using the sampling technique.
For the neighbor sampler, only when the batch size is very small The time spent in the training phase is significantly reduced compared to full-picture training. 
When the batch size is particularly small, the training time is significantly reduced. However, because the sampling itself has additional overhead, 
and there is also an overhead in transferring the sampled sub-image to the GPU, the training time of the entire batch may exceed the full image. 
Training is time-consuming. For cluster sampler, the sampled subgraph is smaller than neighbor sampler under the same relative batch size,
which makes the effect of sampling technology more obvious. But when the relative batch size increases, 
the additional cost of sampling technology The increase is very obvious. When the relative batch size $\geq$ 25\%,
the total time of each batch of the sampling technique even exceeds the time of the full image training. 
Experiments show that \textbf{the implementation of sampling technology in PyG at this stage is very inefficient and the extra overhead is high}.
When the batch size is slightly larger, the additional cost of sampling technology accounts for more than 50\% of the total training cost of the entire batch.
\textbf{Sampling technology can reduce training time only on very small batches}.

>>>>

Fig~\ref{fig:exp_absolute_training_time} compares the training time per epoch for each GNN model, the ranking is GaAN $\gg$ GAT $>$ GGNN $>$ GCN.
Its time-consuming ranking is consistent with the complexity analysis. Because the number of edges in the graph generally far exceeds the number of nodes, 
the GAT with higher edge computing complexity is more time-consuming than the algorithm GGNN with high vertex computational complexity. At the same time, 
Fig~\ref{fig:exp_absolute_training_time} shows that the training time of individual epochs is abnormally high, which is mainly caused by profiling overhead and the GC pause of the Python interpreter.
This phenomenon confirms the necessity of removing abnormal epochs.

@@

\begin{figure}
	\centering
    \subfloat[GCN\label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gcn}]{\includegraphics[height=4cm]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gcn.png}}\\
    \subfloat[GGNN\label{fig:exp_hyperparameter_on_vertex_edge_phase_time_ggnn}]{\includegraphics[height=4cm]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_ggnn.png}}\\
    \caption{The influence of hyperparameters on GCN/GGNN vertex/edge calculation time}
	\label{fig:exp_hyperparameter_on_vertex_edge_phase_time}
\end{figure}

\begin{figure}
	\centering
    \includegraphics[height=8cm]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gat.png}
    \caption{The influence of hyperparameters on GAT vertex/edge calculation time}
	\label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat}
\end{figure}

\begin{figure}
	\centering
    \includegraphics[height=8cm]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gaan.png}
    \caption{The influence of hyperparameters on GaAN vertex/edge calculation time}
	\label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gaan}
\end{figure}

@@

Experiments also show that \textbf{the average degree of the dataset affects the time-consuming proportion of vertex/edge calculation}. 
Our fixed graph has 50k vertices, and the R-MAT generator is used to generate a random graph with an average degree between 10 and 100.
We measure the change of the time-consuming proportion of the vertex/edge calculation in each GNN model varies with the average degree of the graph,
as shown in Fig~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}. The time calculation of the edge increases linearly with the increase of the average degree,
\textbf{The time calculation of the edge calculation dominates the entire calculation time in most cases}. 
Only when the vertex calculation complexity is very high and the average degree is very low, the vertex calculation time can overtake the edge calculation time.
Therefore, \textbf{the focus of GNN training optimization should improve the efficiency of edge calculation}.

@@

\begin{figure}
    \centering
    \includegraphics[width=1\columnwidth]{figs/illustration/steps_in_edge_calculation.png}
    \caption{Step decomposition of edge calculation}
    \label{fig:steps_in_edge_calculation}
\end{figure}

@@

Fig~\ref{fig:exp_sampling_batch_train_time} shows the change of the training time of each batch on the `amc` and `fli` datasets with the batch size after using the sampling technique.
For the neighbor sampler, only when the batch size is very small The time spent in the training phase is significantly reduced compared to full-picture training. 
When the batch size is particularly small, the training time is significantly reduced. However, because the sampling itself has additional overhead, 
and there is also an overhead in transferring the sampled subgraph to the GPU, the training time of the entire batch may exceed the full image. 
Training is time-consuming. For cluster sampler, the sampled subgraph is smaller than neighbor sampler under the same relative batch size,
which makes the effect of sampling technology more obvious. But when the relative batch size increases, 
the additional cost of sampling technology The increase is very obvious. When the relative batch size $\geq$ 25\%,
the total time of each batch of the sampling technique even exceeds the time of the full image training. 
Experiments show that \textbf{the implementation of sampling technology in PyG at this stage is very inefficient and the extra overhead is high}.
When the batch size is slightly larger, the additional cost of sampling technology accounts for more than 50\% of the total training cost of the entire batch.
\textbf{Sampling technology can reduce training time only on very small batches}.
